Molecular transport in human pial perivasculature

### 2024-11-13 paper/supplementary_main.tex
Updates, still WIP.
Former-commit-id: 2f4a6b4897e53e79fdcd9489a6b92ad8b364a5c4 [formerly 2f4a6b4897e53e79fdcd9489a6b92ad8b364a5c4 [formerly bd39ac695ec46dc46e73a39431bad712a1273d11]]
Former-commit-id: 2c0e453cab63dcc64971d3615f8a0264574b7ab4
Former-commit-id: d40b4ce2e1dd33f3ebb5a6f5b16573d9a1175119

In addition, we consider two sets of perivascular networks represented
as graphs: a periarterial network $\Lambda_a$ represented by the
(connected) centerlines $\Lambda^i_{(a)}$ of the arterial tree, and a
perivenous network $\Lambda_v$ associated with the centerlines
$\Lambda^i_{(v)}$ of the veins. Recall that we represent the vascular
domains as the union of cylindrical vessels of radius $R_1^i$
surrounding the centerlines $\Lambda^i$. Moreover, we consider the
surrounding perivascular spaces as the union of annular cylinders
$\Omega^i$ of inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$
and thus of width $R_2^i - R_1^i$. We denote the outer lateral surface
of the periarterial and perivenous spaces by $\Gamma_a$ and
$\Gamma_v$, respectively. We omit the subscript $a$ or $v$ when
referring to any such network, vessel segment or perivascular outer
surface. We assume that $\Lambda$ is parametrized by the coordinate
$s$, and with a minor abuse of notation, simply write $s \in \Lambda$
to represent the point ${\bm \lambda}(s)$ on $\Lambda$ corresponding
to $s$.

@@

In this section, we describe the mathematical
model~\eqref{eq:multi_transport}, the associated definitions,
interface and boundary conditions in further detail. We refer to the
main text for explicit parameter values while giving the general, abstract
form here. Recall that we model a concentration field $c = c(x, t)$
for $x \in \Omega$ and $t > 0$ defined in $\Omega_{\rm PAR}$ and
$\Omega_{\rm CSF}$ separately. In each of these domains, $c$ satisfies
\begin{equation}
  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma = f \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
  \label{eq:3d_pde}
\end{equation}
In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$). In the parenchyma, $\phi$ represents the extracellular space volume fraction. Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$. $\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivacular networks $\Lambda_a, \Lambda_v$. To summarize,
\[
\phi =  \begin{cases}
  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
  \end{cases}, \; 
D = \begin{cases}
  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
  \quad \bm u  = \begin{cases}
  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
\end{cases}, \; 
\xi = \begin{cases}
  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
\end{cases} . 
\]
In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp. $|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{CSF}}$). If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.

Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
\begin{equation}
  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
\end{equation}
where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$. Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the (Dirac delta) term $\delta_\Gamma$ is a distribution concentrated on the outer lateral surfaces of the PVSs. More precisely, it is defined in terms of its \mer{integrated} action on any sufficiently smooth function $v : \Omega \rightarrow \R$ and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$ such that
\begin{equation}
  \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
\end{equation}
under the assumption that $\xi$ is constant on each $\partial
\Theta_2^i$.
%% , since:
%% \begin{multline}
%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
%% \end{multline}
We refer to the reference~\cite{masri2023modelling} for a complete model derivation and analysis.
%% \mer{MER: The previous was:
%% \begin{equation}
%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
%% \end{equation}
%% }

In addtion, we model the concentration $\hat{c} = \hat{c}(s, t)$ in
the periarterial and perivenous networks ($s \in \Lambda_a, s \in
\Lambda_v$) which represents the concentration in the perivascular
space averaged over each perivascular cross-section and satisfies:
\begin{equation}
\partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0  \qquad \quad \mathrm{in} \quad  \Lambda_a, \Lambda_v ,
\end{equation}
where $A = A(s)$ is defined for each $s \in \Lambda^i$ and each
centerline $\Lambda^i$ with PVS $\Omega^i$ as the area of the PVS
cross-section $\Theta^i(s)$.

@@

At bifurcation nodes in $\Gamma_a, \Gamma_v$, we impose flux conservation and continuity of the concentration $\hat c$. We refer to \eqref{eq:flux_conservation} in \Cref{sec:details_numerical_method} for more details. For the perivascular network, the boundary nodes are collected into $\partial \Lambda$. We assume no tracer outflow on all boundary nodes and prescribe a homogeneous Neumann condition: 
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \partial \Lambda. 
\end{equation}

Further, $P$ and $A$ are the perimeter and area of the PVS which can vary along the centerline network, respectively.  Finally, we remark that $\hat u$ takes different values on each $\Lambda_i$.  Both $\Omega$ and $\Lambda$ are fixed in time and we assume constant-in- time radii (despite varying radii to drive peristaltic flow, see \cref{sec:sup:peristalsis}). Finally, we note that the model assumes that there is no interaction between the inner perivascular boundary and the blood vessel.

\paragraph*{Interface conditions}
On the interface $\Gamma_{\mathrm{pia}}$ between the CSF and brain, we prescribe the following condition modeling a semi--permeable pial interface 
\begin{alignat}{2}
(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} &= - (- D \nabla c_{\rm{CSF}} + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n}_{\rm{PAR}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}} ,  \\  
(-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad  \pia \cup \Gamma_{\mathrm{LV}} .  
\end{alignat} 
Here, $\beta_{\mathrm{pia}}$ is a permeability constant between the brain and the CSF:

@@

For the convection term, we use upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the references therein. Define 
\begin{equation*}
a_h^{\mathrm{up}}(c_h, v_h) =- \sum_{E\in\mathcal{T}_h } \int_E \bm{u} c_h \nabla v_h  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c_h\} \cdot \bm{n}_F [v_h]  + \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F |}{2} [c_h ] [v_h ] 
\end{equation*}
The discrete formulation for \eqref{eq:3d_pde} then reads 
\begin{multline}
\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\spinal} \gin v_h , \quad \forall v_h \in V_h^k.
\end{multline}
For the 1D concentration $\hat c$, we recall that $\hat c = \hat c_i $ on each segment of the network $\Lambda^i$.  We enforce continuity of the concentration on the bifurcation points and conservation of fluxes. We let $\mathcal{V}$ denote the set of all vertices in $\Lambda$ and $ 
\mathcal{E}(\mathsf{v})$ denote the set of segments sharing a vertex $\mathsf{v}$.  For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n_i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
n_i(\mathsf{v}_{\mathrm{in}}^i) = 1 , \,\, n _i (\mathsf{v}_{\mathrm{out}}^i) = -1 \,\, \mathrm{and} \,\,\, n_i(\mathsf{v}) = 0, \quad \forall \mathsf{v} \in 
\mathcal{V} \,  \backslash \,  \{ \mathsf{v}_{\mathrm{in}}^\mathsf e,\mathsf{v}_{\mathrm{out}}^\mathsf e \}.  
$$
The collection of bifurcation points is denoted by $\mathcal{B} = \{ \mathsf{v} \in \mathcal{V}, \,\,  \mathrm{card}(\mathcal{E}(\mathsf{v})) \geq  3\}$. For any $\mathsf{v} \in \mathcal{B}$, we have 
\begin{equation}
\sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0  \label{eq:flux_conservation}
\end{equation}
For approximating the solution of the 1D equations in the network, we use the space 
\[ \hat V_h = \{ v_h \in C^0(\Lambda), 
\; v_h \vert_{\Lambda^i} \in V_h(\Lambda^i)
\}, 
\]
where $V_h(\Lambda^i)$ is the space of continuous polynomials of degree $1$ on each $\Lambda^i$,

\subsection{Cardiac Dispersion Factor Estimation}\label{sec:dispersion_details}

>>>>

In addition, we consider two sets of perivascular networks: a
periarterial network $\Lambda_a$ represented by the (connected)
centerlines $\Lambda^i_{(a)}$ of the arterial tree, and a perivenous
network $\Lambda_v$ associated with the centerlines $\Lambda^i_{(v)}$
of the veins. Recall that we represent the vascular domains as the
union of cylindrical vessels of radius $R_1^i$ surrounding the
centerlines $\Lambda^i$. Moreover, we consider the surrounding
perivascular spaces as the union of annular cylinders $\Omega^i$ of
inner radius $R_1^i$ and outer radius $R_2^i > R_1^i$ and thus of
width $R_2^i - R_1^i$. We denote the outer lateral surface of the
periarterial and perivenous spaces by $\Gamma_a$ and $\Gamma_v$,
respectively. We omit the subscript $a$ or $v$ when referring to any
such network, vessel segment or perivascular outer surface. We assume
that $\Lambda$ is parametrized by the coordinate $s$, and with a minor
abuse of notation, simply write $s \in \Lambda$ to represent the point
${\bm \lambda}(s)$ on $\Lambda$ corresponding to $s$. Finally, we view
each perivascular network both as a geometric domain and as a directed
graph with the centerlines $\{ \Lambda^i \}$ as oriented edges and the
connections as nodes $\mathcal{V}$.

@@

In this section, we describe the mathematical
model~\eqref{eq:multi_transport}, the associated definitions,
interface and boundary conditions in further detail. We refer to the
main text for explicit parameter values while giving the general,
abstract form here, and to the reference~\cite{masri2023modelling} for
the derivation and analysis of this 3D-1D model. Recall that we model
a concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
each of these domains, $c$ satisfies
\begin{equation}
  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma = f \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
  \label{eq:3d_pde}
\end{equation}
In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$). In the parenchyma, $\phi$ represents the extracellular space volume fraction. Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$. $\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivacular networks $\Lambda_a, \Lambda_v$. To summarize,
\[
\phi =  \begin{cases}
  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
  \end{cases}, \; 
D = \begin{cases}
  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
  \quad \bm u  = \begin{cases}
  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
\end{cases}, \; 
\xi = \begin{cases}
  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
\end{cases} . 
\]
In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp. $|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{CSF}}$). If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.

Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
\begin{equation}
  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
\end{equation}
where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$. Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the (Dirac delta) term $\delta_\Gamma$ is a distribution concentrated on the outer lateral surfaces of the PVSs. More precisely, it is defined in terms of its \mer{integrated} action on any sufficiently smooth function $v : \Omega \rightarrow \R$ and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$ such that
\begin{equation}
  \mer{\int_{\Omega}} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
\end{equation}
under the assumption that $\xi$ is constant on each $\partial
\Theta_2^i$. 
%% , since:
%% \begin{multline}
%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
%% \end{multline}
%% \mer{MER: The previous was:
%% \begin{equation}
%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
%% \end{equation}
%% }

The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
space averaged over each perivascular cross-section and is governed by 
\begin{equation}
  \partial_t (A  \hat c) - \partial_s(\hat D A \partial_s ( \hat c)) +\partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
\label{eq:1d_pde}
\end{equation}
where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
associated PVS $\Omega^i$, the area of the cross-section
$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
and $\hat{u}$ is a convective velocity representing an (average) CSF
flow velocity in the axial direction of the perivascular spaces. To
complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{T}$, write $\hat{c}_i = \hat{c}|_{\Lambda^i}$, and let
$\mathcal{E}(y)$ denote the set of (one, two or three) edges in
$\{\Lambda^i\}$ sharing the node $y$. To impose continuity of the
concentrations, we set that the concentrations when viewed from each
edge must match at nodes:
\begin{equation}
  \hat{c}_i (y) = \hat{c}_y \qquad \foralls \Lambda^i \in \mathcal{E}(y),
  \quad \foralls y \in \mathcal{B} .
\end{equation}
Moreover, to ensure conservation of flux, we set that the flux going
in and out at nodes should add to zero, or more precisely, that for
each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{i \in \mathcal{E}(y)} (\hat{D} A_i \partial_s \hat{c}_i (y)  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0
  \label{eq:flux_conservation}
\end{equation}

For a given edge $\Lambda^i= (\mathsf{v}_\mathrm{in}^i, \mathsf{v}_{\mathrm{out}}^{i})$, we define the function $n_i: \mathcal{V} \rightarrow \{-1,0,1\}$ with $$
n_i(\mathsf{v}_{\mathrm{in}}^i) = 1 , \,\, n _i (\mathsf{v}_{\mathrm{out}}^i) = -1 \,\, \mathrm{and} \,\,\, n_i(\mathsf{v}) = 0, \quad \forall \mathsf{v} \in 
\mathcal{V} \,  \backslash \,  \{ \mathsf{v}_{\mathrm{in}}^\mathsf e,\mathsf{v}_{\mathrm{out}}^\mathsf e \}.  
$$
The collection of bifurcation points is denoted by $\mathcal{B} = \{ \mathsf{v} \in \mathcal{V}, \,\,  \mathrm{card}(\mathcal{E}(\mathsf{v})) \geq  3\}$. For any $\mathsf{v} \in \mathcal{B}$, we have 
\begin{equation}
\sum_{ i \in \mathcal{E}(\mathsf{v})} (\hat D A_i \phi \partial_s \hat{u}_i(\mathsf{v})  -  A_i \hat u_i(\mathsf{v}) \hat c_i(\mathsf{v})) n_i(\mathsf{v}) = 0  \label{eq:flux_conservation}
\end{equation}

\paragraph{Boundary and interface conditions}

On the interface $\Gamma_{\mathrm{pia}}$ between the CSF and brain, we prescribe the following interface condition modeling a semi--permeable interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and $c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
\begin{subequations}
\begin{alignat}{2}
(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} &= - (- D \nabla c_{\rm{CSF}} + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n}_{\rm{PAR}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}} ,  \\  
(-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad  \pia \cup \Gamma_{\mathrm{LV}} .  
\end{alignat} 
\end{subequations}
Here, $\beta_{\mathrm{pia}}$ is a permeability constant between the
brain and the CSF. \mer{MER: Is the inclusion of $\Gamma_{\rm LV}$
  correct in the equation here? If so, text needs minor edits.}

For the perivascular network, the boundary nodes are collected into
$\partial \Lambda$. We assume no tracer outflow on all boundary nodes
and prescribe a homogeneous Neumann condition:
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \partial \Lambda. 
\end{equation}

%% Both $\Omega$ and $\Lambda$ are fixed in time and we assume
%% constant-in- time radii (despite varying radii to drive peristaltic
%% flow, see \cref{sec:sup:peristalsis}).

@@

\paragraph*{Interface conditions}

@@

For the convection term, we use upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the references therein. Define 
\begin{equation*}
a_h^{\mathrm{up}}(c_h, v_h) =- \sum_{E\in\mathcal{T}_h } \int_E \bm{u} c_h \nabla v_h  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c_h\} \cdot \bm{n}_F [v_h]  + \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F |}{2} [c_h ] [v_h ] 
\end{equation*}
The discrete formulation for \eqref{eq:3d_pde} then reads 
\begin{multline}
\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\spinal} \gin v_h , \quad \forall v_h \in V_h^k.
\end{multline}
For approximating the solution of the 1D equations in the network, we use the space 
\[ \hat V_h = \{ v_h \in C^0(\Lambda), 
\; v_h \vert_{\Lambda^i} \in V_h(\Lambda^i)
\}, 
\]
where $V_h(\Lambda^i)$ is the space of continuous polynomials of degree $1$ on each $\Lambda^i$,

\subsection{Cardiac Dispersion Factor Estimation}

\subsubsection{Cardiac Dispersion Factor Estimation}\label{sec:dispersion_details}
